Introduction to Lattice-based Attacks\\``Cryptography''

### 2021-06-07 lattice_based_attacks.tex
Fix grammars

\usepackage{alltt, fancyvrb, url}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{wrapfig}
% \usepackage{algorithm}% http://ctan.org/pkg/algorithms
% \usepackage[ruled,vlined]{algorithm2e}
% \usepackage{algorithm}
\usepackage[]{algorithm2e}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{fontenc}
\usepackage{amsmath,stmaryrd,mathtools,algorithm}
\usepackage{amssymb}
%\usepackage{amsfonts}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{url}

@@

To introduce the mathematics of lattices a basic linear algebra background is needed, so the second chapter is a summary of the
basics properties needed to understand it. We used as main resource ``An Introduction to Mathematical Cryptography''\cite{mathcrypto14} to
explain them, but we also tried to gives some geometrical intuition behind some operations.

@@

Let $v_1,v_2,\ldots,v_k \in V$. A \textit{linear combination} of $v_1,v_2,\ldots,v_k \in V$ is any vector of the form

@@

\begin{figure}[!tbp]
    \begin{minipage}[b]{0.50\textwidth}
        \includegraphics[width=\textwidth]{./img/lattice_b0.png}
        \caption{Lattice $L$ spanned by $v, w$}
        \label{fig:lattice0}
    \end{minipage}
    \hspace{\fill}
    \hspace{\fill}
    \hspace{\fill}
    \hspace{\fill}
    \hspace{\fill}
    \begin{minipage}[b]{0.50\textwidth}
        \includegraphics[width=\textwidth]{./img/lattice_b1.png}
        \caption{Lattice $L$ spanned by $v', w'$}
        \label{fig:lattice1}
    \end{minipage}
\end{figure}

@@

\texbtf{Example}. Let $L$ a lattice spanned by $v_1 = (0, 1), v_2 = (2, 0)$, and a target vector $t = (2.3, 2.4)$. The closest vector
to $t$ in the lattice $L$ is $w = (2, 2)$.\\

@@

Kelby Ludwig explains an intuition behind every steps of the algorithm \cite{lllintuition}, and Oded Regev
gives a more rigorous explanation of the algorithm and the properties \cite{lllexplanation}.\\

Some of the applications of the algorithm are the following:

@@

\begin{enumerate}
    \item Fix $e = 65537$ or $e = 3$ (public).
    \item Find two primes $p, q$ such that $p - 1$ and $q - 1$ are relatively prime to $e$, i.e. gcd$(e, p-1) = 1$ and gcd$(e, q-1) = 1$.
    \item Compute $N = p * q$ and $\phi(n) = (p-1) * (q-1)$
    \item Calculate $d$ (private) as the multiplicative inverse of $e$ modulo $\phi(n)$.
    \item $(N, e)$ is the public key, $(N, d)$ is the private key.
\end{enumerate}

@@

RSA relies on the hardness of factoring the modulo $N$ and we don't have a polynomial algorithm to factor it, so it's considered secure.
However there are different attacks against textbook RSA or bad implementations:

\begin{itemize}
    \item If $m < N^{\frac{1}{e}}$, then $m^e < N$ and the modulo operation is not applied and we only need to find the $e$th root of $c$ over the
        integers to find $m$.
    \item If $q = p + x$ where $x$ is small enough than it's easy to recover the factors of $N$ computing $A = \sqrt{N}$ and compute
            $p = A - x$ for different value of $x$ until $N \mod p = 0$, then we have found $q = N / p$.
    \item It's possible to recover $d$ if it is too small (for example to accelerate decryption) using the Wiener's attack \cite{wiener}.
    \item Other attacks can be found on the excellent survey by Dan Boneh \cite{boneh99twentyyears}.
\end{itemize}

@@

Suppose $N$ is an \textbf{RSA} modulus and we don't know the factorization of it. Let's have an univariate integer polynomial $f(x)$ with degree $n$

@@

\begin{itemize}
    \item It only does \textbf{integer linear operations} on the basis vectors.
        In this way even if the basis is different it's only a linear combination of vector that still have $x_0$ as root
        modulo $N^m$.
    \item If we craft the lattice properly, the norm of shortest vector on the reduced basis will satisfy \ref{eq:pol}.
        We know the length's bound of the shortest vector that LLL could find.
\end{itemize}

@@

Suppose that we don't know the factorization of $N$ and we would like to know the message $m$.
We know the padding and that the message $m$ is 2 bytes long, composed by only ascii letters, so $m < \texttt{0x7b7b}$.

@@

Thus we have that $c = (a + m)^3 \mod N$, for an unkown small $m$.
We can define $f(x) = (a + x)^3 - c$, and so we setup the problem to find a small root $m$ such that $f(m) \equiv 0 \mod N$

@@

The rows of the matrix correspond to the coefficient vectors of the polynomials $f(x), Nx^2, Nx$ and $N$, furthermore we know
that each polynomials will be 0 modulo $N$ if evaluated at $x = m$.
We applied Howgrave-Graham with $m=1$ (the $N^m$ parameter not the message).\\
With this lattice construction every vector is of the form $v = (v_3X^3, v_2X^2, v_1X, v_0)$, because any
integer linear combination of the vector of the lattice will keep the bound $X^i$ for $i=0,..,\dim(B)-1$.

@@

We have $(\det B)^{\frac{1}{4}} = (X^6N^3)^{\frac{1}{4}} < \frac{N}{\sqrt{4}}$, if we solves for $X$ this will be satisfied when
$X < (\frac{N}{16})^{\frac{1}{6}}$. We can consider the $\sqrt{4}$ an error term to have $X < N^{\frac{1}{6}}$.
With the numbers we have this inequality is verified, however even if the bound of the shortest vector is larger we still have some possibilities to
find the correct root.

If we compute the root of $g(x)$ over the integers we obtain $m = \texttt{0x4142} = $ ``AB'' which is the correct result.\\

@@

We need to verify that the shortest vector that we can with \textbf{LLL} will be less than $p$ because if $g(x) < p$,
then by construction we can compute $g(r) = 0$ over the integers.
In this case we have

@@

The last step is to construct $g(x) = v_2x^2 + v_1x + v_0$ (just divides every coefficient by $X^i$)
and computes the roots over the integers to find $r$. We implemented the attack against a modulus
of \texbtf{RSA}-$1024$ knowing the $160$ most significant bits of $p$. The attack can be expanded
until $X < p^{\frac{1}{2}}$ with a bigger lattice.

@@

\textbf{ROCA}\cite{roca} is a vulnerability discovered in 2017 in a software library \textit{RSALib} in the generation
of the public modulus for \textbf{RSA}. The attack is believed to affects millions of smart cards. The vulnerbality
resides in the generation of prime numbers, since embedded devices don't have lots of computational resources a fast way
was needed. Every prime generated has the following structure:

@@

Where $k, a$ are random values different for every prime (unkown), $g = 65537$ and $M$ it's a \textit{primorial} number.
A \textit{primorial} number $M$ is the product of the the first $n$ successive primes

\[
    M = P_n# = \prod_{i=1}^{n} P_i
\]

@@

The security of \textbf{ECDSA} depends on the discrete logarithm problem.
Given the points $Q, G$ such that $Q = k*G$ it's considered an hard problem to find $k$.
However as RSA there are lots of implementation attacks

@@

The lattice-based attacks presented here are based on the \textit{HNP (Hidden Number Problem)} introduced
by Boneh and Venkatesan \cite{hnp} that was in a second moment generalized to prove that the most significan bits
of $k$ can potentially leak the entire nonce \cite{ecdsa_nonce}.
In 2020 a group of researcher were able to leak the MSBs of \textit{DH} and \textit{DHE} secrets in \textbf{TLS $\le$ 1.2}
and using a similar lattice constructions as the one we will describe to solve the \textit{HNP}
and thus breaks the protocol \cite{cryptoeprint:2020:1151} (Raccoon Attack).

@@

Source of the method:

@@

The results obtained from the implementation of the attacks are satisfactory and can be used against real world practical cases.
There is a fair amount of probability and heuristic to make the attacks works that isn't discussed in detail.
As with other branch of mathematics not all theorems and properties are proved to be useful the moment they were first introduced
but can become useful years later as the hidden number problem that become crucial for the Raccoon Attack.
The study on how to exploit side-channel attacks (ex. knowing some bits of $p$ in \textbf{RSA}) has been proven useful
with \textbf{ROCA}.
The most used cryptographic protocol used has been proved secure but it's very easy to implement them wrong and side-channel
vulnerabilities can heavily compromise the security of the protocol.

>>>>

\usepackage{alltt, fancyvrb, url}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{wrapfig}
% \usepackage{algorithm}% http://ctan.org/pkg/algorithms
% \usepackage[ruled, vlined]{algorithm2e}
% \usepackage{algorithm}
\usepackage[]{algorithm2e}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{fontenc}
\usepackage{amsmath, stmaryrd, mathtools, algorithm}
\usepackage{amssymb}
%\usepackage{amsfonts}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{url}

@@

To introduce the mathematics of lattices a basic linear algebra background is needed, so the second chapter is a summary of the
basics properties needed to understand it. We used as main resource ``An Introduction to Mathematical Cryptography''\cite{mathcrypto14} to
explain them, but we also tried to give some geometrical intuition behind some operations.

@@

Let $v_1,v_2,\ldots,v_k \in V$. A \textit{linear combination} of $v_1,v_2,\ldots,v_k \in V$ is any vector of the form:

@@

\begin{figure}[!bp]
    \begin{minipage}[b]{0.50\textwidth}
        \includegraphics[width=\textwidth]{./img/lattice_b0.png}
        \caption{Lattice $L$ spanned by $v, w$}
        \label{fig:lattice0}
    \end{minipage}
    \hspace{\fill}
    \hspace{\fill}
    \hspace{\fill}
    \hspace{\fill}
    \hspace{\fill}
    \begin{minipage}[b]{0.50\textwidth}
        \includegraphics[width=\textwidth]{./img/lattice_b1.png}
        \caption{Lattice $L$ spanned by $v', w'$}
        \label{fig:lattice1}
    \end{minipage}
\end{figure}

@@

\textbf{Example}. Let $L$ a lattice spanned by $v_1 = (0, 1), v_2 = (2, 0)$, and a target vector $t = (2.3, 2.4)$. The closest vector
to $t$ in the lattice $L$ is $w = (2, 2)$.\\

@@

Kelby Ludwig explains an intuition behind every step of the algorithm \cite{lllintuition}, and Oded Regev
gives a more rigorous explanation of the algorithm and the properties \cite{lllexplanation}.\\

Some applications of the algorithm are the following:

@@

\begin{enumerate}
    \item Fix $e = 65537$ or $e = 3$ (public).
    \item Find two primes $p, q$ such that $p - 1$ and $q - 1$ are relatively prime to $e$, gcd$(e, p-1) = 1$ and gcd$(e, q-1) = 1$.
    \item Compute $N = p * q$ and $\phi(n) = (p-1) * (q-1)$
    \item Calculate $d$ (private) as the multiplicative inverse of $e$ modulo $\phi(n)$.
    \item $(N, e)$ is the public key, $(N, d)$ is the private key.
\end{enumerate}

@@

RSA relies on the hardness of factoring the modulo $N$, and we don't have a polynomial algorithm to factor it, so it's considered secure.
However, there are different attacks against textbook RSA or bad implementations:

\begin{itemize}
    \item If $m < N^{\frac{1}{e}}$, then $m^e < N$ and the modulo operation is not applied and we only need to find the $e$th root of $c$ over the
        integers to find $m$.
    \item If $q = p + x$ where $x$ is small enough than it's easy to recover the factors of $N$ computing $A = \sqrt{N}$ and compute
            $p = A - x$ for different value of $x$ until $N \mod p = 0$, then we have found $q = N / p$.
    \item It's possible to recover $d$ if it is too small (in practice could be used to accelerate decryption) using the Wiener's attack \cite{wiener}.
    \item Other attacks can be found on the excellent survey by Dan Boneh \cite{boneh99twentyyears}.
\end{itemize}

@@

Suppose $N$ is an \textbf{RSA} modulus, and we don't know the factorization of it. Let's have an univariate integer polynomial $f(x)$ with degree $n$

@@

\begin{itemize}
    \item It only does \textbf{integer linear operations} on the basis vectors.
        In this way even if the basis is different it's only a linear combination of vector that still have $x_0$ as root
        modulo $N^m$.
    \item If we craft the lattice properly, the norm of the shortest vector on the reduced basis will satisfy \ref{eq:pol}.
        We know the length's bound of the shortest vector that LLL could find.
\end{itemize}

@@

Suppose that we don't know the factorization of $N$, and we would like to know the message $m$.
We know the padding and that the message $m$ is 2 bytes long, composed by only ascii letters, so $m < \texttt{0x7b7b}$.

@@

Thus, we have that $c = (a + m)^3 \mod N$, for an unkown small $m$.
We can define $f(x) = (a + x)^3 - c$, and setup the problem to find a small root $m$ such that $f(m) \equiv 0 \mod N$

@@

The rows of the matrix correspond to the coefficient vectors of the polynomials $f(x), Nx^2, Nx$ and $N$, furthermore we know
that each polynomial will be 0 modulo $N$ if evaluated at $x = m$.
We applied Howgrave-Graham with $m=1$ (the $N^m$ parameter not the message).\\
With this lattice construction every vector is of the form $v = (v_3X^3, v_2X^2, v_1X, v_0)$, because any
integer linear combination of the vector of the lattice will keep the bound $X^i$ for $i=\dim(B)-1,..,0$.

@@

We have $(\det B)^{\frac{1}{4}} = (X^6N^3)^{\frac{1}{4}} < \frac{N}{\sqrt{4}}$, if we solve for $X$ this will be satisfied when
$X < (\frac{N}{16})^{\frac{1}{6}}$. We can consider the $\sqrt{4}$ an error term to have $X < N^{\frac{1}{6}}$.
With the numbers we have this inequality is verified, however even if the bound of the shortest vector is larger we still have some possibilities to
find the correct root.

Computing the root of $g(x)$ over the integers we obtain $m = \texttt{0x4142} = $ ``AB'' which is the correct result.\\

@@

We need to verify that the shortest vector that we can find with \textbf{LLL} will be less than $p$ because if $g(x) < p$,
then by construction we can compute $g(r) = 0$ over the integers.
In this case we have

@@

The last step is to construct $g(x) = v_2x^2 + v_1x + v_0$ (just divides every coefficient by $X^i$)
and computes the roots over the integers to find $r$. We implemented the attack against a modulus
of \textbf{RSA}-$1024$ knowing the $160$ most significant bits of $p$. The attack can be expanded
until $X < p^{\frac{1}{2}}$ with a bigger lattice.

@@

\textbf{ROCA}\cite{roca} is a vulnerability discovered in 2017 in a software library \textit{RSALib} in the generation
of the public modulus for \textbf{RSA}. The attack is believed to affect millions of smart cards. The vulnerbality
resides in the generation of prime numbers, since embedded devices don't have lots of computational resources a fast way
was needed. Every prime generated has the following structure:

@@

Where $k, a$ are random values different for every prime (unkown), $g = 65537$ and $M$ it's a \textit{primorial} number.
A \textit{primorial} number $M$ is the product of the first $n$ successive primes

\[
    M = P_n\# = \prod_{i=1}^{n} P_i
\]

@@

The security of \textbf{ECDSA} depends on the discrete logarithm problem.
Given the points $Q, G$ such that $Q = k*G$ it's considered an hard problem to find $k$.
However, as RSA, there are lots of implementation attacks

@@

The lattice-based attacks presented here are based on the \textit{HNP (Hidden Number Problem)} introduced
by Boneh and Venkatesan \cite{hnp} that was in a second moment generalized to prove that the most significant bits
of $k$ can potentially leak the entire nonce \cite{ecdsa_nonce}.
In 2020 a group of researcher were able to leak the MSBs of \textit{DH} and \textit{DHE} secrets in \textbf{TLS $\le$ 1.2}
and used a similar lattice constructions as the one we will describe to solve the \textit{HNP}
and thus break the protocol \cite{cryptoeprint:2020:1151} (Raccoon Attack).

@@

@@

The results obtained from the implementation of the attacks are satisfactory and can be used against real world practical cases.
There is a fair amount of probability and heuristic to make the attacks works that isn't discussed in detail.
As with other branch of mathematics not all theorems and properties are proved to be useful the moment they were first introduced
but can become useful years later as the hidden number problem that become crucial for the Raccoon Attack.
The study on how to exploit side-channel attacks (ex. knowing some bits of $p$ in \textbf{RSA}) has been proven useful
with \textbf{ROCA}.
The most used cryptographic protocol used has been proved secure but it's very easy to implement them wrong and side-channel
vulnerabilities can heavily compromise the security of the protocols.
